Digital economy in the UK: a multi-scalar story of the diffusion of web technologies

### 2023-03-21 paper/submission1/tranos2023.tex
paper first draft

---
title: "Digital economy in the UK: a multi-scalar story of the diffusion of web technologies"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol and The Alan Turing Institute
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
  This paper maps the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. The data and geolocation strategy allow to capture these processes at a very granular spatial scale. The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies. Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

# Introduction {#sec:1}

Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. The seminal contribution of
@hagerstrand1968innovation is illustrative of this early interest.
However, the torch of exploring and modelling such processes had been
passed to other disciplines such as economics, business studies and
sociology well before the 'cultural turn' of economic geography
[@perkins2005international]. A potential explanation of lack of studies
exploring the diffusion of new and, more specifically, digital
technologies across *both* space and time can be attributed to the
scarcity of relevant and granular enough data. As @zook2022mapping
highlight, digital activities are hardly ever captured in official data.

This paper aims to reinstate geographers' interest in how new and
digital technologies diffuse over space and time. It adopts a geographic
standpoint to explore and map the diffusion of web technologies in the
UK over space and time. It does so by employing a novel source of big
data which captures the active engagement with web technologies during
the $1996-2013$ period.

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time use this information to make predictions
regarding the diffusion of related future technologies. As per
@meade2021modelling a variety of actors have a direct interest in
gaining such knowledge including network equipment suppliers; network
operators, regulatory and local authorities. And despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a granular level.

@@

@zook2022mapping Because new digital activities are rarely---if
ever---captured in official state data, researchers must rely on
information gathered from alternative sources. With this in mind, the
overarching aim of this article is outlining an approach for analyzing
the "new spaces and geographies" of digital phenomena and practices.

<!-- motivation -->

@@

An a priori approach groups similar countries into segments, where
similarity is defined using economic, social, or political covariates.
Examples include Hofstede (2001) and Lee (1990). An a posteriori
analysis considers segments based on realized market behaviour, for
example see Sood et al. (2009). Some studies investigate changing
segment membership over time, see Cannon and Yaprak (2011).

@fritsch2015new Hägerstrand (1952, 1967) was the first scholar to
intensively analyze spatial diffusion pro cesses. Based on the empirical
observations of the regional spread of process innovations, he
hypothesized that new technology is first implemented in the 'center' --
the large agglom erations -- and then moves 'down' the spatial
hierarchy, the last stop being the 'periphery', i.e. remote and sparsely
populated regions. There are at least two factors that may explain this
pattern: first, the transfer of tacit knowledge via face-to-face contact
and second, the absorptive capacity of a region.

<!-- key concepts -->

@bento2015spatial The concept of absorptive capacity has also been
applied in the TIS context in terms of the con-struction of capacity
needed to transfer and implement a new technology in the receiving
country (cf.Van Alphen, 2011), with repercussions for the nature of
system functions

@bednarz2020pulled larger local technological niches (at least from a
demand perspective) are more likely to be formed in urban areas than in
rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
However, size is not the only regional characteristic that may matter in
this context. Regions also differ in accumulated experiences and the
presence of tacit knowledge with respect to products, as well as in
actors' propensity share this knowledge among producers (Martin et al.,
2019)

<!-- @das2022diffusion -->

<!-- Epidemic approach: The approach emphasizes that the adoption of a -->

<!-- technology spreads like an infection among the potential adopters. By -->

<!-- taking a homogeneous group of potential adopters as an assumption, it -->

<!-- explains the adoption of any innovation. -->

@@

<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->

<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->

<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->

@morrill2020spatial Spatial diffusion is the process by which behavior
or characteristics of the landscape change as a result of what happens
elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
over space and timed, from limited origins

Contagious diffusion=\> spatial autocorelation

@@

<!-- the three geographic diffusion processes are (1) purely contagious where -->

<!-- distance or adjacency is the absolute controlling factor, (2) purely -->

<!-- hierarchical where size or urban position in the central place hierarchy -->

<!-- is the absolute controlling factor, or (3) where the location of change -->

<!-- is purely random. These are the pure or ideal forms of the diffusion -->

<!-- process; all real diffusion processes are the result of a combination of -->

<!-- these extremes. -->

@@

<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->

<!-- are considered as the three pillars of the theory of technological -->

<!-- change and Schumpeter is considered as the father of the study of -->

<!-- diffusion -->

@leibowicz2016representing Diffusion is a spatial (Hagerstrand, 1967) as
well as a temporal process, and historical evidence confirms that
technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies (Victor, 1993).

<!-- @leibowicz2016representing -->

<!-- An illustrative empirically based theory for technology -->

<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->

<!-- (Grubler, 1990). With respect to a specific technology, it divides -->

<!-- regions into three groups: core, rim, and periphery. The core is where -->

<!-- the technology is invented, subject to experimentation, and first -->

<!-- deployed commercially. To develop the technology, the core must have -->

<!-- both sufficient innovative ability and economic motivation to make the -->

<!-- necessary resource commitment. As the technology spreads from the core -->

<!-- to the rim and eventually to the periphery, later adopters have the -->

<!-- advantage of benefiting from the experiences of earlier adopters via -->

<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->

<!-- point, the technology has likely been refined and costs have likely -->

<!-- declined significantly. Schmidt's Law claims that, as a result, -->

<!-- diffusion proceeds at a faster rate in regions which adopt the -->

<!-- technology later. However, diffusion tends to be less pervasive in -->

<!-- regions where adoption is delayed, ultimately saturating at a lower -->

<!-- extent than in regions where adoption occurred earlier. This is often -->

<!-- attributed to a lack of proper institutions and infrastructures in later -->

<!-- adopters, markets whichmay not be as favorable to a technology, or the -->

<!-- availability of yet newer options that invite technological -->

<!-- leapfrogging. -->

@@

wilson201281 Early on in their lifecycle, new technologies are crude,
imperfect, and expensive (Rosenberg, 1994). New energy technologies are
attractive for their ability to perform a particular task or deliver a
new or improved energy service (Fouquet, 2010). This is often
circumscribed by a particular set of needs in a particular context: a
market 'niche'. End-users in niche markets are generally less sensitive
to the effective price of the energy service provided or have a higher
willingness to pay for its performance advantages (Fouquet, 2010). Thus
initially, performance dominates cost competitiveness (Wilson and
Grubler, 2011). Market niches afford some protection from competitive
pressures, allowing technologies to be tested and improved in applied
settings, reducing uncertainties with performance or market demand (Kemp
et al., 1998). Costs may only fall substantively after an extended
period of commercial experimentation, concurrent with the establishment
of an industrial base and characteristic moves towards standardisation
and mass production (Grubler, 1998). The influence of accumulating
production experience on costs is captured by the concept of learning.

<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->

<!-- diffusion is the process by which an innovation is communicated through -->

<!-- certain communication channels over time among the members of a social -->

<!-- system. An innovation may be an idea, object or practice which is -->

<!-- perceived as new by an individual or another unit of adoption (e.g., an -->

<!-- organization) . Communication via communication channels refers to how -->

<!-- knowledge about an innovation is transferred between the members of a -->

<!-- social system. Communication takes time, and is therefore an important -->

<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->

<!-- diffusion studies agree that an innovation does not instantly diffuse -->

<!-- into a social system. Instead a typical time pattern of diffusion along -->

<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->

<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->

<!-- confirmed by a large number of studies· of primitive preindustrial -->

<!-- societies. -->

@@

<!-- grubler1990rise -->

<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->

<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->

<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->

<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->

<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->

<!-- and diffusion eventually ceases. -->

<!-- grubler1990rise -->

<!-- Thus, three main characteristics define the spatial diffusion process: -->

<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->

<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->

<!-- diffusion proceeds from initial centers to secondary ones. Within the -->

<!-- context of urban systems (central places) this implies a hierarchical -->

<!-- spread of an innovation from larger to smaller centers, and from the -->

<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->

<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->

<!-- (Brown, 1981) nearby rather than far-away locations. -->

@grubler1990rise Later Hagerstrand conceptualized physical "barrier"
effects like lakes or uninhabited areas, which, in addition to distance,
act as further retarding effects on diffusion. These are formalized in
the form of "zero" or "half" contact multiplicators on the (distance
decaying) message flows.

@grubler1990rise With respect to the formalization of the communication
flows Hagerstrand defines a "mean information field" (MIF), in which the
probability of communication is a negative function of distance between
individuals

@perkins2005international Theoretical models disagree as to why firms
adopt innovations at different times. Epidemic models emphasize
information (Griliches 1957). Certain firms are hypothesized to adopt
earlier because they come into contact with, and learn from, adopters of
the new technology before others. Economic models, on the other hand,
predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
Firms adopt technologies at different times because they differ with
respect to various organizational and environmental variables
influencing the economic returns from adoption (Blackman 1999).

Either way, the strong assumption is that developing countries can
acquire modern technology innovated in developed economies, often at a
fraction of the original research and development (R&D) costs, thereby
leapfrogging many decades of technological progress (Teece 2000).

The first involves a country's geographical location. Recent empirical
work suggests that diffusion is ''geographically localized'' (Globerman,
Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
diffuses faster in a country where it is already more widely diffused in
neighboring countries. Underlying these regional effects are contagion
and contact with prior users or producers of technology.

@ding2010modeling Rogers (1995) characterized early adopters as
knowledgeable risk takers. Griliches (1957) looked at the rate of return
for early adopters and characterized them as profit maximizers. The
geographer Hagerstrand (1967) was focusing on the character of early
adopters in the transmittal process that produces follow-on participants
in the process. Space was treated as a contiguity system with embedded
characteristics producing barrier elements to the diffusion process.

@ding2010modeling REVIEW ON Technological Diffusion

<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->

<!-- starts from innovative centers ('core´) and disseminates, through a -->

<!-- hierarchy of subcenters, to the periphery. This process is sequential -->

<!-- rather than simultaneous, and tends to accelerate from core to -->

<!-- periphery, where diffusion reaches a lower intensity than in the core -->

<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->

<!-- spatial acceleration has been identified in marketing research and is -->

<!-- known as the "lead-lag effect" (Peres et al., 2010). -->

@perkins2011internet Two main mechanisms are identified in the
literature to explain diffusion: (1) epidemic-type dynamics whereby
contact with previous adopters stimulates uptake as po tential adopters
learn about a new innovation; and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be comesmore profitable,
useful, or valuable, with uptake characteristically spreading as costs
be come lower, performance improves, or the po tential uses of the
innovation grow over time.

@@

<!-- methods -->

@wilson201281 Logistic growth describes an initial period of gradual
diffusion as a technology is introduced as a new commercial application,
moving then through a rapid, exponential growth phase, before slowing
and eventually saturating [@grubler1999dynamics]. The substitution of
incumbent technologies by new competitors leads to subsequent decline
and eventual obsolescence.

# Literature review {#sec2} MAYBE SKIP THIS??

<!-- previous studies -->

@beardsell1999spatial To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992

@@

@bednarz2020pulled spatial innovation diffusion By using Bayesian
survival models with time-dependent data of wind turbine deployment and
firm foundation for 402 German regions between the years 1970 and 2015,
we show that the spatial evolution of the German wind energy industry
was more strongly influenced by local demand-- pull than local
supply--push processes.

@@

@perkins2005international analyze whether the rate at which new producer
technologies diffuse is significantly influenced by (1) latecomer
advantage and (2) engagement with the global economy via trade and
foreign investment

@@

@fritsch2015new analyze the spatial diffusion of laser technology
research in West Germany from 1960, when this technology began, until
2005.

@perkins2011internet

@lengyel2020role We see that the OSN spread almost exclusively from the
original location (the capital Budapest, with an order of magnitude more
inhabitants than the next size town) to various parts of the country in
the early phase of the life-cycle. Later, diffusion became less
mono-centric and other towns also emerged as spreaders. Our findings
support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in
Figure 2I.

@PAPAGIANNIDIS2015308

@leibowicz2016representing for energy There is no generally accepted
theory that explains diffusion rate heterogeneity across technologies,
but several factors are considered important. Greater unit scale and
larger market size contribute to slower diffusion. Requirements for
interrelated technologies or complex infrastructures also hinder the
diffusion process (Grubler, 2012).

@@

@leibowicz2016representing Empirical evidence supports the validity of
Schmidt's Law over a wide range of technologies, time periods, and
geographical contexts. A recent meta-analysis of technology up-scaling
found that diffusion accelerated moving from the core to the rim and
periphery for technologies as diverse as natural gas power, oil
refineries, and automobiles (Wilson, 2009). One historical example that
conforms particularly well to Schmidt's Law is the diffusion of coal
power in Europe (Grubler, 2012). England emerged as the core region for
coal power because it had legal and economic institutions that
incentivized scientific pursuits, domestic coal reserves, and a clear
industrial motivation to replace water power with coal. ...

@ding2010modeling model the spatial diffusion of mobile
telecommunications in China as well as its determinants. regional
socioeconomic characteristics play an important role in determining the
timing, speed, and level of mobile telecommunications diffusion in China

@bento2018time explore What determines the duration of formative phases
for energy innovations in different markets? We are interested both in
initial markets (also: core, lead, first mover, early adopter) where
formative phases prepare technologies for mass commercialization, and in
follower markets (also: periphery, lag, late adopter) where accelerated
formative phases may benefit from diffusion and spillovers.

# Materials and Methods {#sec3}

@wilson201281 growth function description in p. 86

from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson
k/(1 + e\^(t_0 - t)/scal)

Wilson: k(1 + e\^-b(t - t_0)) So, scal = - 1/b

The curve is symmetric\* around t_0

**minimum R2 = 95%**, see also @grubler1990rise

@@

Moran's I as @ding2010modeling **TODO** for t_0, diffusion speed

## Results {#sec3.1}

@@

**TODO** maps at the local authority level per firms and discuss
patterns

Neighbourhood effect: diffusion proceeds outwards from innovation
centers, first "hitting" nearby rather than far-away locations (Grubler
1990)

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Digital economy in the UK: a multi-scalar story of the diffusion
of web technologies}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

\affiliation[1]{organization={University of Bristol and The Alan Turing
Institute},postcode={UK},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper maps the participation in the digital economy and its
evolution in the UK over space and time. Most of the existing economic
geography literature which dealt with the spatiality of the internet
employed supply-side measures, such as infrastructural capacity, in
order to understand the geography of the digital economy and its
potential spatial economic effects. Useful as these approaches might
have been, they cannot capture the micro-processes and the
characteristics of the individual online behaviour. Using large volumes
of archived and geolocated web content, this paper models the diffusion
of web technologies over space and time in the UK. The data and
geolocation strategy allow to capture these processes at a very granular
spatial scale. The modelling approach, which is based on simple spatial
analytical methods and on the estimation of diffusion curves at various
scales, enables to depict the role of geography and other cognitive
factors which drove the diffusion of web technologies. Although the
focus is on a recent historical period -- 1996-2012 -- the results of
the analysis depict diffusion mechanisms which can be very useful in
understanding the evolutionary patterns of the adoption of other newer
technologies.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. The seminal contribution of
\citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of lack of
studies exploring the diffusion of new and, more specifically, digital
technologies across \emph{both} space and time can be attributed to the
scarcity of relevant and granular enough data. As
\citet{zook2022mapping} highlight, digital activities are hardly ever
captured in official data.

This paper aims to reinstate geographers' interest in how new and
digital technologies diffuse over space and time. It adopts a geographic
standpoint to explore and map the diffusion of web technologies in the
UK over space and time. It does so by employing a novel source of big
data which captures the active engagement with web technologies during
the \(1996-2013\) period.

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time use this information to make predictions
regarding the diffusion of related future technologies. As per
\citet{meade2021modelling} a variety of actors have a direct interest in
gaining such knowledge including network equipment suppliers; network
operators, regulatory and local authorities. And despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a granular level.

@@

\citet{zook2022mapping} Because new digital activities are rarely---if
ever---captured in official state data, researchers must rely on
information gathered from alternative sources. With this in mind, the
overarching aim of this article is outlining an approach for analyzing
the ``new spaces and geographies'' of digital phenomena and practices.

@@

An a priori approach groups similar countries into segments, where
similarity is defined using economic, social, or political covariates.
Examples include Hofstede (2001) and Lee (1990). An a posteriori
analysis considers segments based on realized market behaviour, for
example see Sood et al.~(2009). Some studies investigate changing
segment membership over time, see Cannon and Yaprak (2011).

\citet{fritsch2015new} Hägerstrand (1952, 1967) was the first scholar to
intensively analyze spatial diffusion pro cesses. Based on the empirical
observations of the regional spread of process innovations, he
hypothesized that new technology is first implemented in the `center' --
the large agglom erations -- and then moves `down' the spatial
hierarchy, the last stop being the `periphery', i.e.~remote and sparsely
populated regions. There are at least two factors that may explain this
pattern: first, the transfer of tacit knowledge via face-to-face contact
and second, the absorptive capacity of a region.

\citet{bento2015spatial} The concept of absorptive capacity has also
been applied in the TIS context in terms of the con-struction of
capacity needed to transfer and implement a new technology in the
receiving country (cf.Van Alphen, 2011), with repercussions for the
nature of system functions

\citet{bednarz2020pulled} larger local technological niches (at least
from a demand perspective) are more likely to be formed in urban areas
than in rural ones, which is in line with the work of Ha¨gerstrand
(1965a,b). However, size is not the only regional characteristic that
may matter in this context. Regions also differ in accumulated
experiences and the presence of tacit knowledge with respect to
products, as well as in actors' propensity share this knowledge among
producers (Martin et al., 2019)

@@

\citet{morrill2020spatial} Spatial diffusion is the process by which
behavior or characteristics of the landscape change as a result of what
happens elsewhere earlier. Spatial diffusion is the spread of the
phenomenon, over space and timed, from limited origins

Contagious diffusion=\textgreater{} spatial autocorelation

@@

@@

\citet{leibowicz2016representing} Diffusion is a spatial (Hagerstrand,
1967) as well as a temporal process, and historical evidence confirms
that technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies (Victor, 1993).

@@

wilson201281 Early on in their lifecycle, new technologies are crude,
imperfect, and expensive (Rosenberg, 1994). New energy technologies are
attractive for their ability to perform a particular task or deliver a
new or improved energy service (Fouquet, 2010). This is often
circumscribed by a particular set of needs in a particular context: a
market `niche'. End-users in niche markets are generally less sensitive
to the effective price of the energy service provided or have a higher
willingness to pay for its performance advantages (Fouquet, 2010). Thus
initially, performance dominates cost competitiveness (Wilson and
Grubler, 2011). Market niches afford some protection from competitive
pressures, allowing technologies to be tested and improved in applied
settings, reducing uncertainties with performance or market demand (Kemp
et al., 1998). Costs may only fall substantively after an extended
period of commercial experimentation, concurrent with the establishment
of an industrial base and characteristic moves towards standardisation
and mass production (Grubler, 1998). The influence of accumulating
production experience on costs is captured by the concept of learning.

@@

\citet{grubler1990rise} Later Hagerstrand conceptualized physical
``barrier'' effects like lakes or uninhabited areas, which, in addition
to distance, act as further retarding effects on diffusion. These are
formalized in the form of ``zero'' or ``half'' contact multiplicators on
the (distance decaying) message flows.

\citet{grubler1990rise} With respect to the formalization of the
communication flows Hagerstrand defines a ``mean information field''
(MIF), in which the probability of communication is a negative function
of distance between individuals

\citet{perkins2005international} Theoretical models disagree as to why
firms adopt innovations at different times. Epidemic models emphasize
information (Griliches 1957). Certain firms are hypothesized to adopt
earlier because they come into contact with, and learn from, adopters of
the new technology before others. Economic models, on the other hand,
predominantly emphasize firm heterogeneity (Ireland and Stoneman 1986).
Firms adopt technologies at different times because they differ with
respect to various organizational and environmental variables
influencing the economic returns from adoption (Blackman 1999).

Either way, the strong assumption is that developing countries can
acquire modern technology innovated in developed economies, often at a
fraction of the original research and development (R\&D) costs, thereby
leapfrogging many decades of technological progress (Teece 2000).

The first involves a country's geographical location. Recent empirical
work suggests that diffusion is `'geographically localized'' (Globerman,
Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
diffuses faster in a country where it is already more widely diffused in
neighboring countries. Underlying these regional effects are contagion
and contact with prior users or producers of technology.

\citet{ding2010modeling} Rogers (1995) characterized early adopters as
knowledgeable risk takers. Griliches (1957) looked at the rate of return
for early adopters and characterized them as profit maximizers. The
geographer Hagerstrand (1967) was focusing on the character of early
adopters in the transmittal process that produces follow-on participants
in the process. Space was treated as a contiguity system with embedded
characteristics producing barrier elements to the diffusion process.

\citet{ding2010modeling} REVIEW ON Technological Diffusion

\citet{perkins2011internet} Two main mechanisms are identified in the
literature to explain diffusion: (1) epidemic-type dynamics whereby
contact with previous adopters stimulates uptake as po tential adopters
learn about a new innovation; and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be comesmore profitable,
useful, or valuable, with uptake characteristically spreading as costs
be come lower, performance improves, or the po tential uses of the
innovation grow over time.

@@

\citet{wilson201281} Logistic growth describes an initial period of
gradual diffusion as a technology is introduced as a new commercial
application, moving then through a rapid, exponential growth phase,
before slowing and eventually saturating \citep{grubler1999dynamics}.
The substitution of incumbent technologies by new competitors leads to
subsequent decline and eventual obsolescence.

\hypertarget{literature-review-sec2-maybe-skip-this}{%
\section{Literature review \{\#sec2\} MAYBE SKIP
THIS??}\label{literature-review-sec2-maybe-skip-this}}

\citet{beardsell1999spatial} To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992

@@

\citet{bednarz2020pulled} spatial innovation diffusion By using Bayesian
survival models with time-dependent data of wind turbine deployment and
firm foundation for 402 German regions between the years 1970 and 2015,
we show that the spatial evolution of the German wind energy industry
was more strongly influenced by local demand-- pull than local
supply--push processes.

@@

\citet{perkins2005international} analyze whether the rate at which new
producer technologies diffuse is significantly influenced by (1)
latecomer advantage and (2) engagement with the global economy via trade
and foreign investment

@@

\citet{fritsch2015new} analyze the spatial diffusion of laser technology
research in West Germany from 1960, when this technology began, until
2005.

\citet{perkins2011internet}

\citet{lengyel2020role} We see that the OSN spread almost exclusively
from the original location (the capital Budapest, with an order of
magnitude more inhabitants than the next size town) to various parts of
the country in the early phase of the life-cycle. Later, diffusion
became less mono-centric and other towns also emerged as spreaders. Our
findings support the idea that spreading initially happens to large
distances and becomes more local over time. This is illustrated and
discussed later in Figure 2I.

\citet{PAPAGIANNIDIS2015308}

\citet{leibowicz2016representing} for energy There is no generally
accepted theory that explains diffusion rate heterogeneity across
technologies, but several factors are considered important. Greater unit
scale and larger market size contribute to slower diffusion.
Requirements for interrelated technologies or complex infrastructures
also hinder the diffusion process (Grubler, 2012).

@@

\citet{leibowicz2016representing} Empirical evidence supports the
validity of Schmidt's Law over a wide range of technologies, time
periods, and geographical contexts. A recent meta-analysis of technology
up-scaling found that diffusion accelerated moving from the core to the
rim and periphery for technologies as diverse as natural gas power, oil
refineries, and automobiles (Wilson, 2009). One historical example that
conforms particularly well to Schmidt's Law is the diffusion of coal
power in Europe (Grubler, 2012). England emerged as the core region for
coal power because it had legal and economic institutions that
incentivized scientific pursuits, domestic coal reserves, and a clear
industrial motivation to replace water power with coal. \ldots{}

\citet{ding2010modeling} model the spatial diffusion of mobile
telecommunications in China as well as its determinants. regional
socioeconomic characteristics play an important role in determining the
timing, speed, and level of mobile telecommunications diffusion in China

\citet{bento2018time} explore What determines the duration of formative
phases for energy innovations in different markets? We are interested
both in initial markets (also: core, lead, first mover, early adopter)
where formative phases prepare technologies for mass commercialization,
and in follower markets (also: periphery, lag, late adopter) where
accelerated formative phases may benefit from diffusion and spillovers.

\hypertarget{sec3}{%
\section{Materials and Methods}\label{sec3}}

\citet{wilson201281} growth function description in p.~86

from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
+ e\^{}(t\_0 - t)/scal)

Wilson: k(1 + e\^{}-b(t - t\_0)) So, scal = - 1/b

The curve is symmetric* around t\_0

\textbf{minimum R2 = 95\%}, see also \citet{grubler1990rise}

@@

Moran's I as \citet{ding2010modeling} \textbf{TODO} for t\_0, diffusion
speed

\hypertarget{sec3.1}{%
\subsection{Results}\label{sec3.1}}

@@

\textbf{TODO} maps at the local authority level per firms and discuss
patterns

Neighbourhood effect: diffusion proceeds outwards from innovation
centers, first ``hitting'' nearby rather than far-away locations
(Grubler 1990)
